---
title: Text, audience, accessibility
format: gfm
---

```{r}
#| label: setup
# source formatters script. add any other scripts in your utils folder you have
source(here::here("utils/formatters.R"))
library(ggplot2)
```

## Adding text within a chart

Go back to lab 2 where we defined a function to place labels slightly inward from the end of a bar. Copy and paste that code here. 

```{r}
#| label: label-nudge-function
# placeholder---this just returns the input value
label_nudge <- function(x) {
    x
}
```

Add code to the definition of dental_visit to add 2 new variabled: `nudged_value`, a nudged position for the labels, and `pct_value`, where you've formatted the labels as percentages. Do this either in 1 or 2 `dplyr::mutate` calls. You might want to rewrite the `percent` function that's in the formatters script to fit a different number of decimal places (`accuracy` argument).

If the label placement doesn't look the way you want, add or adjust arguments to your `label_nudge` function until you like the way things are calculated.

```{r}
#| label: label-nudge-1
dental_visit <- justviz::cdc |>
    dplyr::filter(
        location %in%
            c(
                "US",
                "Maryland",
                "Baltimore city",
                "Baltimore County",
                "Howard County",
                "Anne Arundel County"
            )
    ) |>
    dplyr::filter(indicator == "Dental visit") |>
    dplyr::mutate(location = forcats::as_factor(location)) |>
    dplyr::mutate(value = value / 100) |>
    # placeholders---change with your own code
    dplyr::mutate(nudged_value = label_nudge(value)) |>
    dplyr::mutate(pct_value = value)

ggplot(dental_visit, aes(x = location, y = value)) +
    geom_col() +
    geom_text(aes(y = nudged_value, label = pct_value), size = 5)
```

Does that function work as well when the chart is flipped to horizontal? Adjust arguments of the function and in your `geom_text` call.

```{r}
#| label: label-nudge-2
ggplot(dental_visit, aes(x = value, y = location)) +
    geom_col() +
    geom_text(aes(x = nudged_value, label = pct_value), size = 5) +
    # something I learned this week is that this works for discrete scales!
    coord_cartesian(reverse = "y")
```

## Audience

Copy over the code for one of your lab 9 charts with your nice theme applied. Imagine you're a consultant to a local government agency (pick one depending on what data is in your chart). Make 2 revisions of your chart:

- One version you'd use to present your research to a meeting of officials from the agency. You can use whatever type of title you want, but it should be at no higher than a 10th grade reading level.
- One version you'd suggest they use for a public awareness campaign on the issue, such as social media posts, handouts at clinics or government offices, posters on bus shelters, etc. Use narrative titles ("Baltimore County has a higher...") at no higher than a 7th grade reading level.

When you design for each audience, consider your chart type and your use of color, text, and annotations.

Use [wordcalc](https://www.wordcalc.com/readability/) or another online Flesch-Kincaid grade level checker. (On wordcalc, you want the score that says "Flesch-Kincaid Grade Level").


For both charts, write alt text that will explain the chart to someone using a screen reader (put it in the `fig-alt` comment). Follow the recommendations from [Amy Cesal's post](https://medium.com/nightingale/writing-alt-text-for-data-visualization-2a218ef43f81?sk=32db60d651933b5ac2c5b6507f3763b5).


```{r}
#| label: chart-agency
#| fig-alt: ""
#
```


```{r}
#| label: chart-public
#| fig-alt: ""
#
```

By the end of this lab, you should have charts that are close to what will get you a good grade on your midterm project!