---
title: Exploratory data visualization
format: gfm
---

For this lab, you'll do some EDV with the ACS data, taking similar steps to what I did in the EDV course notes. Start with a subset of just data for tracts, and some variables you'll explore. Choose at least 6 numeric variables, plus keep county and name. (Don't use ages 25+, poverty status determined, or total housing units; those are all just there as denominators.)

```{r}
#| label: setup
library(ggplot2)
test_acs <- justviz::acs

acs_tracts <- justviz::acs |>
    dplyr::filter(level == "tract") |>
    dplyr::select(county, name, median_hh_income, no_vehicle_hh, poverty, low_income, diversity_idx)
```

Look at the first few rows of data, and print a summary. Jot down anything that seems noteworthy---missing values, very high or low values, etc.

Nothing noteworthy, correct number of unique counties, very few NAs for each.

```{r}
#| label: summary
summary(acs_tracts)
```

## Variation

For each numeric variable in your data, make a histogram. Adjust the number of bins or the binwidths for each one until you find something easy to read and that shows the data well. Again, jot down what you notice about the distributions. 

They all follow distributions with nothing that visually stand out to me, except for median_hh_income, but that was explained in class: the highest bin is so large because ACS just assigns any tracts with a median hh income > \$250,000 as \$250,001.
No vehicle HH and poverty rate both have extreme skews.

```{r}
#| label: histograms
library(patchwork)
median_hh_income <- ggplot(acs_tracts, aes(x = median_hh_income)) +
    geom_histogram(color = "white")
no_vehicle_hh <- ggplot(acs_tracts, aes(x = no_vehicle_hh)) +
    geom_histogram(color = "white", binwidth = 0.01)
poverty <- ggplot(acs_tracts, aes(x = poverty)) +
    geom_histogram(color = "white")
low_income <- ggplot(acs_tracts, aes(x = low_income)) +
    geom_histogram(color = "white")
diversity_idx <- ggplot(acs_tracts, aes(x = diversity_idx)) +
    geom_histogram(color = "white")
median_hh_income + no_vehicle_hh /
  poverty + low_income + diversity_idx
```

## Unusual values

For one of your variables with a heavy skew or some extreme values, filter the data either with `dplyr::filter` or by setting axis limits like I did in the notes. What can you see by removing these extreme values? 

```{r}
#| label: extremes
acs_tracts |>
    dplyr::filter(no_vehicle_hh < 0.5) |>
    ggplot(aes(x = no_vehicle_hh)) +
    geom_histogram(color = "white", binwidth = .01)
acs_tracts |>
    dplyr::filter(poverty < 0.5) |>
    ggplot(aes(x = poverty)) +
    geom_histogram(color = "white", binwidth = .01)
```


## Covariation

Next, choose one of your numeric variables and make a series of boxplots of it by county. To do this, you can put the variable you're studying on the x-axis and county on the y-axis, then use `geom_boxplot`. What patterns can you see now that were obscured by looking at all tracts in the state lumped together?

```{r}
#| label: boxplots
ggplot(acs_tracts, aes(x = no_vehicle_hh, y = county)) +
  geom_boxplot()
```

Pick a variable you want to investigate; pretend you're going to build a model to predict this variable (dependent variable). Choose another variable that you think could be a feature in your model (independent variable), and make a scatterplot with your dependent variable on the y-axis and your independent variable on the x-axis. If it's too dense to read easily, try different strategies to reduce overplotting. Repeat this with 2 more independent variables.

```{r}
#| label: scatter-1
ggplot(acs_tracts, aes(y = no_vehicle_hh, x = median_hh_income)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = lm)
```

```{r}
#| label: scatter-2
ggplot(acs_tracts, aes(y = no_vehicle_hh, x = poverty)) +
  geom_point()
```

```{r}
#| label: scatter-3
ggplot(acs_tracts, aes(y = no_vehicle_hh, x = diversity_idx)) +
  geom_point()
```

Now pick one of those independent variables that you think could potentially be used in a linear regression model. On your scatterplot, add a regression line with `geom_smooth(method = lm)` (see [the docs](https://ggplot2.tidyverse.org/reference/geom_smooth.html))

```{r}
#| label: scatter-lm
ggplot(acs_tracts, aes(y = no_vehicle_hh, x = poverty)) +
  geom_point() +
  geom_smooth(method = lm)
```

What does the regression line tell you?

It tells me that the data do indeed follow a linear regression but there is a lot of variation in the low values
