---
title: Tables
format: html
tbl-cap-location: top
---

This will be a very quick lab where you just practice formatting your data and making a table of it. Not much that isn't already in the notes.

- Use the ACS data, filtered for the US, Maryland, and a few counties of your choice. 
- Choose a few columns you're interested in, and select those (plus location). 
- Use more than one type of numeric variable (choose a mix of percentages, integers, and/or dollar amounts).
- Define the formatter functions you'll need, and use them to format each variable as appropriate.
- Create a table using `knitr::kable`.
- Give your columns clean, readable names.
- Align the columns so text columns are aligned on the left, numeric ones on the right.
- Give the table a title using the `caption` argument of `kable`.


```{r}
#| label: setup
library(dplyr)
library(justviz)
library(knitr)
library(scales)

# subsetting the dataset to the USA, MD, and the counties around Anne Arundel. 
# Selecting focus variables
acs_anne_around <- justviz::acs |>
    dplyr::filter(
        name %in%
            c(
                "United States",
                "Maryland",
                "Anne Arundel County",
                "Calvert County",
                "Prince George's County",
                "Howard County",
                "Baltimore County",
                "Baltimore city"
            )
    ) |>
    dplyr::select(level, name, median_hh_income, no_vehicle_hh, poverty, low_income, diversity_idx)
```


```{r}
#| label: table-with-kable

# setting up scales functions

percent <- scales::label_percent(accuracy = 1)
dollar <- scales::label_currency(accuracy = 1)

# assigning new values from scales functions
acs_anne_around_fmt <- acs_anne_around |> 
  mutate(median_hh_income = dollar(median_hh_income),
         no_vehicle_hh = percent(no_vehicle_hh),
         poverty = percent(poverty),
         low_income = percent(low_income),
         diversity_idx = percent(diversity_idx)
         )

kable(acs_anne_around_fmt, col.names = c(
  "Level",
  "Name",
  "Median Household Income",
  "No Vehicle Households",
  "Poverty Rate",
  "Low Income Rate",
  "Diversity Index"
),
align = "llrrrrr",
caption = "Selected socioeconomic indicators, Anne Arundel and nearby counties, 2024")
```